On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-03-30 manuscripts/PNAS/chazthm_pnas.tex
getting closer

Here, we prove very generally that uniqueness in sparse linear coding is an expected property of the model.  In other words, whenever enough sparsely represented data generated from a dictionary (satisfying a spark condition) are observed, the original codes and dictionary are uniquely determined up to an error that is commensurate with the measurement noise.  Applications of this result include:  (1) a sufficient diversity of sparse codes sent through a random linear compressive channel are unique, (2) the sparse linear coding problem is well-posed in the sense of Hadamard \cite{Hadamard1902}, and (3) an explanation for universal representation of natural signals in neuroscience.

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $E \subseteq {[m] \choose k}$ with the SIP, there are $N =  |E| \left[ 1 + (k-1){m \choose k} \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}
\begin{proof}
Something quick here.
\end{proof}

%\begin{proposition}
%If $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
%\end{proposition}
Note that if $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property, then $|E| \geq \ell m/k$.
This is easy to see from $|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m$. In certain cases, there exist simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging the elements of $[m]$ in a square grid.  In light of Cor.~\ref{DeterministicUniquenessCorollary}, it would be  interesting problem to determine minimal such $E$.

There are other less direct consequences of Thm.~\ref{DeterministicUniquenessTheorem} to stability in sparse linear coding.  The following implication relates the result to more common optimization formulations of the problem.

\begin{theorem}
Fix integers $n, \ell$ and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k+1)\left[ 1 + (k-1){ m \choose k} \right]$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ satisfying the SIP. There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every solution to the constrained $\ell_0$ optimization problem:
\begin{align}\label{minsum}
\arg \min_{ \substack{ \mathbf{B} \in \mathbb{R}^{n \times m} \\ \mathbf{\hat x}_i, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m} } } \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ \text{s.t.} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{$i \in [N]$},
\end{align}
necessarily satisfies \eqref{Cstable}.
\end{theorem}

\begin{proof}
[*** Incomplete ***]
We argue for a lower bound on the number of $k$-sparse vectors $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. Let $n_p$ be the number of vectors $\mathbf{\hat x}_i$ satisfying $\|\mathbf{\hat x}_i\|_0 = p$, so that $\sum_{p = 1}^{\hat m} n_p = N$. Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
\begin{align}
k \sum_{p = 1}^{m} n_p = kN \geq \sum_{i=1}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=1}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=1}^{m} p n_p 
\end{align}
\[\implies \sum_{p \leq k} (k-p)n_p \geq \sum_{p > k} (p-k) n_p.\]
Hence,
\begin{align}
k \sum_{p \leq k} n_p \geq \sum_{p \leq k} (k-p)n_p \geq \sum_{p > k} (p-k) n_p \geq \sum_{p > k} n_p
\end{align}
\[ = N - \sum_{p \leq k} n_p \implies \sum_{p \leq k} n_p \geq N / (k+1).\]

@@

Another straightforward application of Thm.~\ref{DeterministicUniquenessTheorem} is a probabilistic extension that uses the following fact in random matrix theory.  A random matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} with probability one (or with ``high probability'' for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
in which $\gamma$ is a positive constant that depends on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. \cite[Sec.~4]{Baraniuk08}).

@@

[ ** TODO ** elaborate a little on these? ]

@@

Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} and suppose that for some $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ ($\hat m \geq m$) and $1$-sparse $\mathbf{\hat x}_i \in \mathbb{R}^{\hat m}$ we have  $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon < L_2(\mathbf{A}) / C_1$ for all $i$. Then there are $c'_1, \ldots, c'_m \in \mathbb{R}$ and $\pi: [m] \to [\hat m]$ with:
\begin{align}\label{1D}
\|c_j\mathbf{A}_j - c'_j\mathbf{B}_{\pi(j)}\|_2 \leq \varepsilon,\ \ \   \text{for all $j$}.
\end{align} 
Note that $c'_j \neq 0$ for all $j$ since otherwise by \eqref{delrho} and definition of $C_1$ we would have $\|c_j\mathbf{A}_j\|_2 < \min_{\ell \in [m]}\|c_{\ell}\mathbf{A}_{\ell}\|_2$.

We  now show that $\pi$ is injective (and thus is a permutation if $\hat m = m$). Suppose that $\pi(i) = \pi(j) = \ell$ for some $i \neq j$ and $\ell$. Then, $\|c_{j}\mathbf{A}_{j} - c'_{j}\mathbf{B}_{\ell}\|_2 \leq \varepsilon$ and $\|c_{i}\mathbf{A}_{i} - c'_{i} \mathbf{B}_{\ell}\|_2  \leq \varepsilon$. Scaling and summing these inequalities by $|c'_{i}|$ and $|c'_{j}|$, respectively, and applying the triangle inequality, we obtain:
\begin{align*}%\label{contra}
(|c'_{i}| + |c'_{j}|) \varepsilon
&\geq\|\mathbf{A}(c'_{i}c_{j} \mathbf{e}_{j} - c'_{j}c_{i}\mathbf{e}_{i})\|_2 \nonumber \\ 
&\geq  \left( |c'_{i}| + |c'_{j}| \right) L_2(\mathbf{A}) \min_{\ell \in [m]} |c_\ell |,
\end{align*}
%
which contradicts the bound $\varepsilon < L_2(\mathbf{A})/C_1$. Hence, the map $\pi$ is injective. Setting $J = \pi([m])$ and letting $P = \left( \mathbf{e}_{\pi(1)} \cdots \mathbf{e}_{\pi(m)}\right)$ and $D = \text{diag}(\frac{\hat{c}_1}{c_1},\ldots,\frac{\hat{c}_m}{c_m})$, we see that \eqref{1D} becomes, for all $j \in [m]$:
\begin{align*}%\label{k=1result}
\|\mathbf{A}_j - (\mathbf{B}_J\mathbf{PD})_j\|_2 
= \|\mathbf{A}_j - \frac{c'_j}{c_j}\mathbf{B}_{\pi(j)}\|_2 
\leq \frac{\varepsilon}{|c_j|} 
\leq C_1\varepsilon.
\end{align*}
\end{proof}

@@

Our result in combinatorial matrix analysis is the following, which we derive by the following arguments (see the Supplemental for the full proof). First, we show that the aforementioned proximity between $k$-dimensional subspaces implies a proximity between smaller subspaces spanned by columns of $\mathbf{A}$ indexed by the intersections of sets in $E$ and those spanned by as many or fewer columns of $\mathbf{B}$. Another pigeonholing argument here combined with our assumptions on $E$ (e.g., the singleton intersection property) reveals that, in fact, each column of $\mathbf{A}$ spans a subspace proximal to that spanned by some column of $\mathbf{B}$. \eqref{Cstable} is a simple consequence of this last fact, which actually constitutes the proof of the theorem for the case $k=1$. % We present this special case now before restating the above arguments in greater detail.

\begin{lemma}\label{MainLemma}
Fix integers $n$ and $k < m$, and suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition}. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$. If for some  $\mathbf{B} \in \mathbb{R}^{n \times m}$ and regular $E \subseteq2^{[m]}$ satisfying the singleton intersection property there exists a size-preserving map $\pi: E \mapsto 2^{[m]}$ satisfying:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon,\ \ \   \text{for all $S \in E$},
\end{align}
%
then there exists some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
\begin{align}\label{MainLemmaBPD}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq C_2 \varepsilon, \ \ \  \text{for } j \in [m].
\end{align}
\end{lemma}

@@

%%% No |S| = k assumption %%%
We begin by showing that for every $S \in E$ there is some $S' \subseteq [m]$, $|S'| \leq k$ for which the distance $d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_S'\})$ is controlled by $\varepsilon$. Fix $S \in E$. Since there are $(k-1){m \choose k}+1$ vectors $\mathbf{x}_i$ supported on $S$, the pigeon-hole principle implies that there exists some $S' \in {[m] \choose k}$ and some set of $k$ indices $J(S)$ such that the supports of all $\mathbf{x}_i$ and $\mathbf{\hat x}_i$ with $i \in J(S)$ are contained in $S$ and $S'$, respectively.

Let $\mathbf{X}$ and $\mathbf{X}'$ be the $m \times N$ matrices with columns $\mathbf{x}_i$ and $\mathbf{\hat x}_i$, respectively. It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{J(S)}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_{J(S)}$ are linearly independent and thus form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there then exists a unique $\mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_{J(S)}\mathbf{c}$. Letting \mbox{$\mathbf{y'} = \mathbf{BX}'_{J(S)}\mathbf{c}$}, which is in $\text{\rm Span}\{\mathbf{B}_{S'}\}$, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{y'}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_{J(S)} - \mathbf{BX}'_{J(S)})_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
%\leq \frac{\varepsilon \sqrt{k}}{L(\mathbf{A}X_{J(S)})} \|\mathbf{A}X_{J(S)}\mathbf{c}\|_2 
= \frac{\varepsilon}{L_k(\mathbf{AX}_{J(S)})} \|\mathbf{y}\|_2,
\end{align*}
where the last equality is by definition of the domain-restricted matrix lower bound $L_k$. It follows that for all $S \in E$,
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{S'}\}) 
\leq \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }.
%\leq \frac{\varepsilon}{L(\mathbf{AX}_{J(S)})}.
%&< \left( \frac{ L_2(\mathbf{A}) }{ 2^{|E|} \max_{j \in [m]} \|\mathbf{A}_j\|_2 } \right) \frac{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } { L(\mathbf{AX}_{J(S)}) } \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) \\
\end{align}
The result then follows by  Lem.~\ref{MainLemma}, with the map $\pi: E \to {[m] \choose k}$ defined by the association $S \mapsto S'$ above.
%\begin{align}
%\delta= \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } = \frac{ C \varepsilon }{ \tilde C } < \frac{L_2(\mathbf{A})}{\tilde C}
%\end{align}
\end{proof}

@@

 In this note, we generalize the approach of \cite{Hillar15} to the uniqueness of solutions to Prob.~\ref{InverseProblem} for noisy measurements while also reducing the number of required samples.
% from $N=k{m \choose k}^2$ to $N = m(k-1){m \choose k}+m$. 
% CUT OK? Surprisingly, almost all $n \times m$ dictionaries satisfying the standard assumption \eqref{CScondition} from compressed sensing (CS) are identifiable from enough generic noisy $k$-sparse linear combinations of their elements, up to an error linear in the noise. Moreover, if solutions are constrained to satisfy \eqref{SparkCondition}, then only an upper bound on the number of dictionary elements need be taken as given. 
We remark that our result in the deterministic case (Thm.~\ref{DeterministicUniquenessTheorem}) accounts for \emph{worst-case} noise, whereas the ``effective'' noise might be much smaller when it is sampled from a given distribution; in such cases, the constants $C_1, C_3$ in our theorems %Thms.~\ref{DeterministicUniquenessTheorem},~\ref{DeterministicUniquenessTheorem2} 
will be much smaller with high probability.
We note also that these results extend trivially to cases where point-wise injective nonlinearities are applied to the data. 
We close with several application areas.

@@

>>>>

Here, we prove very generally that uniqueness in sparse linear coding is an expected property of the model.  In other words, whenever enough sparsely represented data generated from a dictionary (satisfying a spark condition) are observed, the original codes and dictionary are uniquely determined up to an error that is commensurate with the measurement noise.  Applications of this result include:  1) a sufficient diversity of sparse codes sent through a random linear compressive channel are unique, 2) the sparse linear coding problem is well-posed in the sense of Hadamard \cite{Hadamard1902}, and 3) an explanation for universal representation of natural signals in neuroscience.

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $E \subseteq {[m] \choose k}$ with the SIP, there are $N =  |E| \left[ 1 + (k-1){m \choose k} \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}
\begin{proof}
Given Thm.\ref{DeterministicUniquenessTheorem}, one needs only to produce sparse codes in general linear position.  However, this is straightforward (e.g., with a ``Vandermonde'' construction \cite{Hillar15}).
\end{proof}

%\begin{proposition}
%If $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
%\end{proposition}
As mentioned above, it is easy to construct such $E$ with $|E| = m$, implying the lower bound for sample size $N$ from the introduction.
Note that if $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the SIP, then $|E| \geq \ell m/k$: \[|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m.\] In certain cases, there are simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging $[m]$ in a square grid.  In light of Cor.~\ref{DeterministicUniquenessCorollary}, it would be interesting to determine other minimal such $E$.

There are other less direct consequences of Thm.~\ref{DeterministicUniquenessTheorem} to stability in sparse linear coding.  The following implication relates the result to more common optimization formulations.

\begin{theorem}
Fix integers $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k+1)\left[ 1 + (k-1){ m \choose k} \right]$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ satisfying the SIP. There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every solution $B$ to the constrained $\ell_0$ optimization:
\begin{align}\label{minsum}
\arg \min_{ \substack{ \mathbf{B} \in \mathbb{R}^{n \times m} \\ \mathbf{\hat x}_i, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m} } } \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ \text{s.t.} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{$i \in [N]$},
\end{align}
necessarily satisfies \eqref{Cstable}.
\end{theorem}

\begin{proof}
We argue for a lower bound on the number of $k$-sparse vectors $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
[*** Incomplete ***]
Let $n_p$ be the number of vectors $\mathbf{\hat x}_i$ satisfying $\|\mathbf{\hat x}_i\|_0 = p$, so that $\sum_{p = 1}^{\hat m} n_p = N$. Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
\begin{align}
k \sum_{p = 1}^{m} n_p = kN \geq \sum_{i=1}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=1}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=1}^{m} p n_p 
\end{align}
\[\implies \sum_{p \leq k} (k-p)n_p \geq \sum_{p > k} (p-k) n_p.\]
Hence,
\begin{align}
k \sum_{p \leq k} n_p \geq \sum_{p \leq k} (k-p)n_p \geq \sum_{p > k} (p-k) n_p \geq \sum_{p > k} n_p
\end{align}
\[ = N - \sum_{p \leq k} n_p \implies \sum_{p \leq k} n_p \geq N / (k+1).\]

@@

Another straightforward application of Thm.~\ref{DeterministicUniquenessTheorem} is a probabilistic extension that uses the following often used application of random matrix theory to compressed sensing.  A random matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} with probability one (or with ``high probability'' for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
in which $\gamma$ is a positive constant that depends on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. \cite[Sec.~4]{Baraniuk08}).

@@

% [ ** TODO ** elaborate a little on these? ]

@@

Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} and suppose that for some $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ ($\hat m \geq m$) and $1$-sparse $\mathbf{\hat x}_i \in \mathbb{R}^{\hat m}$ we have  $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon < L_2(\mathbf{A}) / C_1$ for all $i$. Then there are $\hat{c}_1, \ldots, \hat{c}_m \in \mathbb{R}$ and $\pi: [m] \to [\hat m]$ with:
\begin{align}\label{1D}
\|c_j\mathbf{A}_j - \hat{c}_j\mathbf{B}_{\pi(j)}\|_2 \leq \varepsilon,\ \ \   \text{for all $j$}.
\end{align} 
Note that $\hat{c}_j \neq 0$ for all $j$ since otherwise by \eqref{delrho} and definition of $C_1$ we would have $\|c_j\mathbf{A}_j\|_2 < \min_{\ell \in [m]}\|c_{\ell}\mathbf{A}_{\ell}\|_2$.

We  now show that $\pi$ is injective (and thus is a permutation if $\hat m = m$). Suppose that $\pi(i) = \pi(j) = \ell$ for some $i \neq j$ and $\ell$. Then, $\|c_{j}\mathbf{A}_{j} - \hat{c}_{j}\mathbf{B}_{\ell}\|_2 \leq \varepsilon$ and $\|c_{i}\mathbf{A}_{i} - \hat{c}_{i} \mathbf{B}_{\ell}\|_2  \leq \varepsilon$. Scaling and summing these inequalities by $|\hat{c}_{i}|$ and $|\hat{c}_{j}|$, respectively, and applying the triangle inequality, we obtain:
\begin{align*}%\label{contra}
(|\hat{c}_{i}| + |\hat{c}_{j}|) \varepsilon
&\geq\|\mathbf{A}(\hat{c}_{i}c_{j} \mathbf{e}_{j} - \hat{c}_{j}c_{i}\mathbf{e}_{i})\|_2 \nonumber \\ 
&\geq  \left( |\hat{c}_{i}| + |\hat{c}_{j}| \right) L_2(\mathbf{A}) \min_{\ell \in [m]} |c_\ell |,
\end{align*}
%
which contradicts the bound $\varepsilon < L_2(\mathbf{A})/C_1$. Hence, the map $\pi$ is injective. Setting $J = \pi([m])$ and letting $P = \left( \mathbf{e}_{\pi(1)} \cdots \mathbf{e}_{\pi(m)}\right)$ and $D = \text{diag}(\frac{\hat{c}_1}{c_1},\ldots,\frac{\hat{c}_m}{c_m})$, we see that \eqref{1D} becomes, for all $j \in [m]$:
\begin{align*}%\label{k=1result}
\|\mathbf{A}_j - (\mathbf{B}_J\mathbf{PD})_j\|_2 
= \|\mathbf{A}_j - \frac{\hat{c}_j}{c_j}\mathbf{B}_{\pi(j)}\|_2 
\leq \frac{\varepsilon}{|c_j|} 
\leq C_1\varepsilon.
\end{align*}
\end{proof}

@@

Our result in combinatorial matrix analysis is the following.

%, which we derive by the following arguments (see the Supplemental for the full proof). First, we show that the aforementioned proximity between $k$-dimensional subspaces implies a proximity between smaller subspaces spanned by columns of $\mathbf{A}$ indexed by the intersections of sets in $E$ and those spanned by as many or fewer columns of $\mathbf{B}$. Another pigeonholing argument here combined with our assumptions on $E$ (e.g., the singleton intersection property) reveals that, in fact, each column of $\mathbf{A}$ spans a subspace proximal to that spanned by some column of $\mathbf{B}$. \eqref{Cstable} is a simple consequence of this last fact, which actually constitutes the proof of the theorem for the case $k=1$. % We present this special case now before restating the above arguments in greater detail.

\begin{lemma}\label{MainLemma}
Fix integers $n$ and $k < m$, and suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition}. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$. If for some  $\mathbf{B} \in \mathbb{R}^{n \times m}$ and regular $E \subseteq2^{[m]}$ satisfying the singleton intersection property there exists a size-preserving map $\pi: E \mapsto 2^{[m]}$ satisfying:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon,\ \ \   \text{for all $S \in E$},
\end{align}
%
then there exists a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ with:
\begin{align}\label{MainLemmaBPD}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq C_2 \varepsilon, \ \ \  \text{for } j \in [m].
\end{align}
\end{lemma}

@@

%%% No |S| = k assumption %%%
We begin by showing that for every $S \in E$ there is some $S' \subseteq [m]$, $|S'| \leq k$ for which the distance $d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_S'\})$ is controlled by $\varepsilon$. Fix $S \in E$. Since there are $(k-1){m \choose k}+1$ vectors $\mathbf{x}_i$ supported on $S$, the pigeon-hole principle implies that there exists some $S' \in {[m] \choose k}$ and some set of $k$ indices $I(S)$ such that the supports of all $\mathbf{x}_i$ and $\mathbf{\hat x}_i$ with $i \in I(S)$ are contained in $S$ and $S'$, respectively.

Let $\mathbf{X}$ and $\mathbf{\hat{X}}$ be the $m \times N$ matrices with columns $\mathbf{x}_i$ and $\mathbf{\hat x}_i$, respectively. It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{J(S)}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_{J(S)}$ are linearly independent and thus form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there then exists a unique $\mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_{J(S)}\mathbf{c}$. Letting \mbox{$\mathbf{\hat{y}} = \mathbf{B\hat{X}}_{J(S)}\mathbf{c}$}, which is in $\text{\rm Span}\{\mathbf{B}_{S'}\}$, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\hat{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_{J(S)} - \mathbf{B\hat{X}}_{J(S)})_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
%\leq \frac{\varepsilon \sqrt{k}}{L(\mathbf{A}X_{J(S)})} \|\mathbf{A}X_{J(S)}\mathbf{c}\|_2 
= \frac{\varepsilon}{L_k(\mathbf{AX}_{J(S)})} \|\mathbf{y}\|_2,
\end{align*}
where the last equality is by definition of the domain-restricted matrix lower bound $L_k$. It follows that for all $S \in E$,
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{S'}\}) 
\leq \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }.
%\leq \frac{\varepsilon}{L(\mathbf{AX}_{J(S)})}.
%&< \left( \frac{ L_2(\mathbf{A}) }{ 2^{|E|} \max_{j \in [m]} \|\mathbf{A}_j\|_2 } \right) \frac{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } { L(\mathbf{AX}_{J(S)}) } \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) \\
\end{align}
The result then follows by  Lem.~\ref{MainLemma}, with the map $\pi: E \to {[m] \choose k}$ defined by the association $S \mapsto S'$ above.
%\begin{align}
%\delta= \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } = \frac{ C \varepsilon }{ \tilde C } < \frac{L_2(\mathbf{A})}{\tilde C}
%\end{align}
\end{proof}

@@

In this note, we generalize the approach of \cite{Hillar15} for showing uniqueness in Prob.~\ref{InverseProblem} to the case of noisy measurements while also reducing the sufficient number of samples.
% from $N=k{m \choose k}^2$ to $N = m(k-1){m \choose k}+m$. 
% CUT OK? Surprisingly, almost all $n \times m$ dictionaries satisfying the standard assumption \eqref{CScondition} from compressed sensing (CS) are identifiable from enough generic noisy $k$-sparse linear combinations of their elements, up to an error linear in the noise. Moreover, if solutions are constrained to satisfy \eqref{SparkCondition}, then only an upper bound on the number of dictionary elements need be taken as given. 
We remark that our result in the deterministic case (Thm.~\ref{DeterministicUniquenessTheorem}) accounts for \emph{worst-case} noise, whereas the ``effective'' noise might be much smaller when it is sampled from a given distribution; in such cases, the constants $C_1, C_3$ in our theorems %Thms.~\ref{DeterministicUniquenessTheorem},~\ref{DeterministicUniquenessTheorem2} 
will be smaller. %We note also that these results extend trivially to cases where point-wise injective nonlinearities are applied to the data. 
We close with several application areas.

@@

As a final remark, we hope that researchers in sparse linear coding continue to improve on the bounds and constants here.
